---
title: "Data Tidying"
author: "KMD"
date: "January 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(dplyr)
library(tidyr)
```
If there are errors after library calls, it's because there are functions with the same name in multiple packages.  The last package that was called will overlay previous ones or base R.  For example, when dplyr is loaded in, filter and lag are overlaid from the stats package.  Loading stats package after dplyr would mask the dplyr versions with the stats version.

To call a function from a specific package (in particular ones that are masked), the syntax is: `package_name::function_name(...)`

# Data Cleaning

Read in a data file

```{r}
catch <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)

head(catch)
```

Ctrl+Shift+m gives %>% (pipe operator)


##Learning how to gather

Move from wide to long format data:

```{r}

catch_long <- catch %>% 
  select(-All, -notesRegCode) %>% 
  gather(key = "Species", value = "Catch", -Year, -Region)

head(catch_long)


```

THe code in English: "Catch_long gets the data frame catch, which then is selected to remove the columns All and notesRegCode, then is gathered so that we have the information for species and catch.  Then (after figuring out what went wrong) we changed  the bad value to an integer (and forced it to be an integer as opposed to a character), and multiplied catch_thousands by 1000 to get the total number of the catch"



```{r}
catch_cleaned <- catch_long %>% 
  rename(Catch_thousands = Catch) %>% 
  mutate(Catch_thousands = ifelse(Catch_thousands == "I", 1, Catch_thousands)) %>% 
  mutate(Catch_thousands = as.integer(Catch_thousands))  %>% 
  mutate(Catch = Catch_thousands*1000)
  

head(catch_cleaned)
```

```{r, eval=F, echo=F}
#This is the testing that we did to figure out what had gone wrong
#... the mutate lines wouldn't run because there is a letter/character in the catch_thousands column.  So we try to force it to be an integer (below), which also throws an error because R doesn't know how to deal with the character.  So, we force the character to be an na, and find it using the which function (line 65).  And then we can identify what the problem is and fix it.
test_catch <- as.integer(catch_cleaned$Catch_thousands)

i <- which(is.na(test_catch == TRUE)) #returns the rows with na

catch_cleaned[i,]
```


# Split-Apply-Combine

Calculate total catch by region

```{r}
catch_total <- catch_cleaned %>% 
  group_by(Region) %>% 
  summarize(catch_region = sum(Catch), 
            n_obs=n())

catch_total

```

And you can also group by multiple variables, such as region & year

```{r}
catch_total_RegYear <- catch_cleaned %>% 
  group_by(Region, Year) %>% 
  summarize(catch_region = sum(Catch))

catch_total_RegYear
```
Matt's challenge:  Mean catch by species each year

```{r}
catch_species_Year <- catch_cleaned %>% 
  group_by(Species, Year) %>% 
  summarize(species_mean = mean(Catch),
            n_obs=n())

catch_species_Year
```

Filter for Chinook salmon:
```{r}
catch_chinook_chum <- catch_cleaned %>%
  filter(Species == "Chinook" | Species =="Chum")

head(catch_chinook_chum)
```

Can also stack multiple filters with multiple operators.  `Arrange` allows you to change the order of the output.  In this case the `-Year` arranges the rows in descending order (must add - because ascending order is the default)

```{r}
catch_chinook_SSE_after1990 <- catch_cleaned %>%
  filter(Species == "Chinook" & Region =="SSE" & Year > 1990) %>% 
  arrange(-Year)

head(catch_chinook_SSE_after1990)
```


# Joins

Import data:
```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)

head(region_defs)

```

Clean up the data set to only have the columns we are interested in (region code and name of management area)
```{r}
region_clean <- region_defs %>% 
  select(code, mgmtArea)

head(region_clean)
```

Now we join!

Left joins are almost always used.  `left_join(left_data, right_data, by = "L-key" = "R-key")`

```{r}
catch_joined <- left_join(catch_cleaned, region_clean, 
                          by = c("Region" = "code"))

head(catch_joined)
```

## Spread

Make a wide data frame using `spread`

```{r}
catch_wide <- catch_cleaned %>% 
  filter(Year > 1990) %>% 
  select(-Catch_thousands) %>% 
  spread(key = Year, value = Catch)

head(catch_wide)
```


# Separate & Unite

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df

```
Separate the single date column into a month, day, and year column
```{r}
dates_sep <- dates_df %>% 
  separate(col=date, into=c("month", "day", "year"), sep="/")

head(dates_sep)
```

Now unite into a date the way we want it to be... YYYY-MM-DD.  Still need to figure out how to get it into a 2-digit month.

```{r}
dates_unite <- dates_sep %>% 
  unite(date_iso, year, month, day, sep="-")

head(dates_unite)
```


