---
title: "Untitled"
author: "KMD"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F, echo=F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
```
# Data Import

Read in data from archive.  Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. https://doi.org/10.5063/F1S46Q6M.

```{r, echo=F}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"),
                    stringsAsFactors = FALSE)

head(esc)
```

# Data Tidying

Calculate annual, regional total escapement.

* `separate` the date into year-month-day
* `group_by` and `summarize` to calculate the sum over unique region-year-species combinations
* `filter` for salmon species only

```{r, echo=F}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(SASAP.Region, Species, year) %>% 
  summarise(escapement=sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Pink", "Chum", "Coho"))

```

Creating a dynamic table using the DT package 

```{r}
datatable(annual_esc)
```


# Static Plots



# Nifty Maps!
