---
title: "Untitled"
author: "KMD"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F, echo=F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
```
# Data Import

Read in data from archive.  Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. https://doi.org/10.5063/F1S46Q6M.

```{r, echo=F}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"),
                    stringsAsFactors = FALSE)

head(esc)
```

# Data Tidying

Calculate annual, regional total escapement.

* `separate` the date into year-month-day
* `group_by` and `summarize` to calculate the sum over unique region-year-species combinations
* `filter` for salmon species only

```{r, echo=F}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(SASAP.Region, Species, year) %>% 
  summarise(escapement=sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Pink", "Chum", "Coho"))

```

Creating a dynamic table using the DT package 

```{r}
datatable(annual_esc)
```


# Static Plots

Pretty plots in ggplot2

`aes()` is the aesthetic call and maps everything in your plot to a column in your dataset

*To change things like color, do it in your plot-type call (`geom_col()` in this case)

```{r}
ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement)) +
  geom_col(fill = "blue")
```

But if you want to make colors different based on your data, you CAN put the `fill` in the `aes()` call:

```{r}
ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement, fill = Species)) +
  geom_col()
```



```{r}
ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col()
```

You can also make other neat plots using filters (filtered to JUST the Kodiak data).


```{r}
ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = escapement, color=Species)) +
  geom_line()
```
We switched to `color` from `fill` because the former is generally used for lines and the latter for shapes.

We can also add multiple geometries to a single plot:
```{r}
ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = escapement, color=Species)) +
  geom_line()+
  geom_point()+
  ylab("Escapement (number of fish)")+
  xlab("Year")+
  ggtitle("Kodiak Salmon Escapement")+
  theme_bw()
```

If you want to keep up with the same sort of plots, then you can set a theme as an object and then can call back to it rather than type it all out again.

## Facet Wrap




Use the ggplot 2 [cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)!

# Nifty Maps!



