---
title: "ROpenSci Demo"
author: "KMD"
date: "January 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ROpenSci Demonstration

## Using NOAA data

Requires the use of the "rnoaa" and "ncdf4" packages.

```{r}
library(rnoaa)
library(ncdf4)
library(ggplot2)
```

Read in buoy data:

```{r}
bd <- buoy(dataset = "cwind", buoyid = 46080, year = 2018, datatype = "c")
```

```{r}
wind_speed <- data.frame(time = strptime(bd$data$time, format = "%Y-%m-%dT%TZ"), speed = bd$data$wind_spd, stringsAsFactors = F)
```
*`as.POSIXct()` tells R that this is a date or time, and allows you to do math on it.  Even though it didn't work this time because there is a T and Z.

*So instead we used `strptime()` to reformat the data/time format.

```{r}
ggplot(wind_speed, aes(x = time, y = speed))+
  geom_point(color = "blue")+
  theme_bw()+
  ylab("Wind Speed m/s")+
  xlab("Date")
```

# Now for Mapr

Load libraries

```{r}
library(mapr)
library(spocc)
```

###Plot locations of GBIF species occurrence data for grizzly, polar, and SPIRIT bears

Load in the data!

```{r}
spp <- c("Ursus arctos", "Ursus maritimus", "Ursus americanus kermodei")

bears <- occ(query = spp,
             from = "gbif",
             has_coords = T,
             limit = 500)
```

Plot them up!

```{r}
map_leaflet(bears, color=c("black", "red", "green"), size = 1, alpha = 1)
```

